---
title: "Lab06NotTeamOne"
author: "David Orozco, Arie del Valle, Ryan Tate, Anderson Mun, Ethan Schacht"
date: "February 20, 2019"
output: html_document
---
```{r, echo=FALSE, include=FALSE}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")

```
****
### Team Section
****

```{r}

```
****

### Individual Section
****
```{r}

```
****

#### David's Section
****
```{r, echo = FALSE, message=FALSE}
decflight<-COflights %>% filter(MONTH == "12", !is.na(ARR_DELAY),!is.na(DEP_DELAY))
holiday <- COflights %>% filter(MONTH == "12", DAY_OF_MONTH <= 27, DAY_OF_MONTH >= 23, !is.na(ARR_DELAY),!is.na(DEP_DELAY))

ggplot(data = decflight,mapping = aes(x = ARR_DELAY, y = DEP_DELAY))+
  geom_point() + geom_smooth() + ggtitle('December Delays') + xlab('Arival Delay') + ylab('Departure Delay')
  
```

* I used 'filter' to seperate all the December flights in 2017. 

* There is a strong positive coorelation between between arrival and departure delay.

****
```{r, echo=FALSE}
A <- decflight %>% filter(ARR_DELAY >= 15,DEP_DELAY >= 15) %>% count()
B <- holiday %>% filter(ARR_DELAY >= 15,DEP_DELAY >= 15) %>% count()
C <- (B/A)*100
C
```

* I used 'filter' to seperate all the flights in decemeber that had a delay larger than 15 min. I did the same thing to a tibble of all flights during 23rd-27th of December only. 

* There is a 29.1% probabilty that your flight would have been delayed 15min or more in December of 2017 given you flew during the holidays (23rd - 27th of Dec).

****
#### Arie's Section
****
```{r}

```

****
#### Ethan's Section
****
```{r}

```

****
#### Ryan's Section
****
```{r}

```

****
#### Anderson's Section
****
```{r}

```